---
title: "ps_4.Rmd"
author: "Celia Concannon"
date: "2/22/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
library(tidyverse)
library(gt)
library(dplyr)

orig <- read_csv(
  "ps_4_elections-poll-nc09-3.csv",
                 col_types = cols(
                   .default = col_character(),
                   turnout_scale = col_double(),
                   turnout_score = col_double(),
                   w_LV = col_double(),
                   w_RV = col_double(),
                   final_weight = col_double(),
                   timestamp = col_datetime(format = "")))
# We use col_types like in class, the code would still run without it but would produce an annoying message
```


```{r, echo=FALSE}
dem <- orig %>%
  count(response) %>%
  filter(response == "Dem")
# Count records the number of responses for each category, filter narrows this down to just democrats which is what we're looking for
```

```{r, echo=FALSE}
diff <- orig %>%
  count(response) %>%
  spread(response, n) %>%
  mutate(rep_und = Rep - Und)
# Here I again count the occurences of each response
# Spread creates a table that displays each category of responses with the number of occurences
# Next I used mutate to create a new variable rep_und which is formed by subtracting the number of Undecided from Republicans
```

```{r, echo=FALSE}
diffval <- orig %>%
  filter(gender != gender_combined) %>%
  nrow()
# filter selects for all the individuals where gender is not equal to gender_combined
# nrow counts the number of these individuals
```


```{r, echo=FALSE}
white <- orig %>%
  filter(race_eth == "White") %>%
  filter(file_race_black != "White") %>%
  nrow()
# the first filter selects for all the instances where race_eth is listed as White and then piped into the next filter which selects within that for all the instances where file_race_black is not listed as White
```

```{r, echo=FALSE}
rep_time <- orig %>% 
  filter(response == "Rep") %>%
  select(timestamp) %>%
  arrange(timestamp) %>%
  slice(1)

dem_time <- orig %>%
  filter(response == "Dem") %>%
  select(timestamp) %>%
  arrange(timestamp) %>%
  slice(1)

time_difference <- as.numeric(rep_time - dem_time)

# Here I created two variables the same way but by just filtering for different responses.  Then using Select, arrange, and slice(1), the first response of each is isolated.
# Next I created time_difference by subtracting the two variables
```

There were `r dem$n[1]` respondents who supported the Democratic candidate.

There were `r diff$rep_und[1]` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r diffval` individuals for whom these variables have different values.

There are `r white` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Rep came `r round(time_difference)`  minutes (rounded to the nearest minute) before the first response of Dem.



```{r make_chart, echo=FALSE, results="asis", warning=FALSE}
orig %>%
  select(final_weight, response, race_eth) %>%
# These are the only variables we care about for the purpose of the graph
  mutate(race_eth = fct_relevel(race_eth, c("White", "Black", "Hispanic", "Asian", "Other"))) %>%
# fct_relevel orders each race the way it is shown in the template graph
  group_by(race_eth, response) %>%
  summarize(total = sum(final_weight)) %>%
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%
# This gets ride of the [DO NOT READ]s which we dont want in our data
  spread(key = response, value = total, fill = 0) %>%
# Creates the table
  ungroup() %>%
  mutate(all = Dem + Rep + Und + `3`) %>%
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>%
  mutate(Und = Und / all) %>%
# Created a new variable called "all" which is the sum of all responses from each category
# Used new variable all to create values we will later turn into percentages
  select(-all, -`3`) %>%
  na_if(0) %>%
# Gets rid of the weird value that is returned under Undecided-Hispanic
  gt() %>%
    tab_header(
      title = "Polling Results in North Carolina 9th Congressional District") %>%
    tab_source_note(
      source_note = "Source: New York Times Upshot/Siena Data") %>%
    cols_label(
      race_eth = "Race",
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND.") %>%
    fmt_percent(columns = vars(Dem, Rep, Und),
                decimals = 0)
# gt() allows us to finalize the graph now that we have the values in place, changing title, and header names
# fmt_percent turns our values into percentages and adds the percent sign
# The decimal = 0 argument is so that there are only whole number percentages in the graph
```


```{r, echo=FALSE, warning=FALSE}
orig %>%
  select(educ, final_weight) %>%
# These are the two variables we will be graphing together
  filter(educ != "[DO NOT READ] Refused") %>%
# Gets rid of the [DO NOT READ]s, as we did before
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
# Reorders the categories from educ on the y-axis as they are shown in the template graph
  ggplot(aes(x = educ, y = final_weight)) +
  geom_violin() +
  coord_flip() +
# Flips the x and y axis, without this function the graph doesn't look like a violin plot
  geom_jitter(size = 1, alpha = 0.5, height = 0) +
# Jitter adds the points as seen below, and the arguments within allow their appearance to be played with until I found something that resembled the template graph more closely
  labs(
    title = "More Educated Matter Less in North Carolina 9th",
    subtitle = "Poll gives more weight to people who are less likely to participate in polls",
    source = "New York Times Upshot/Siena College 2018 live polls"
  ) +
  ylab("Weight Given to Respondent in Calculating Poll Results") +
  xlab("")
# Labs names the graph and axis
```


```{r, echo=FALSE, warning=FALSE}
orig %>%
  filter(ager != "[DO NOT READ] Refused") %>%
  filter(partyid != "[DO NOT READ] Refused") %>%
  filter(response != "3") %>%
  ggplot(aes(x = ager, y = response, color = partyid)) +
  geom_jitter() +
  labs(x = NULL) +
  coord_flip() +
  theme(legend.position = "top") +
  scale_color_manual(breaks = c("Democrat", "Republican", "Independent (No party)", "or as a member of another political party"),
                        values=c("blue", "yellow", "green", "red")) +
  labs(
    title = "How Registered Party Members Responded Broken Down by Age",
    source = "New York Times Upshot/Siena College 2018 live polls"
  ) +
  xlab("Age Group") +
  ylab("") +
  theme(legend.title=element_blank())

# Filtered ager and partid to take out [DO NOT READ] Refused because that is irrelevant to the graph
# After reading Data Visualization I chose to do a geom_jitter plot
# I had to play around with scale_color_manual because I wanted those who responded Democrat to be represented by a blue dot and Republicans by a red dot since those are their typical colors
# Labs added title, source, and axis labels
# theme(legend.title = element_blank()) removed the legend title to clean up the graph
# I wanted my graph to show how party registration affected responses, and how this was affected by age
# From my graph it looks like older democrats are more likely to respond Republican
```

## Colleagues: Tate Green